Edge effects in Alberta

### 2021-07-10 Manuscript/manuscript.Rmd
Gave up on TMB
Doesn't work well enough for large numbers of samples

---
title: "Edge effects and optimal field sizes in Alberta"
author:
  - name: Samuel V. J. Robinson
    email: samuel.robinson@ucalgary.ca
    affiliation: University of Calgary
    footnote: 1
  - name: Lan H. Nguyen
    email: hoanglan.nguyen@ucalgary.ca
    affiliation: University of Calgary
  - name: Paul Galpern
    email: paul.galpern@ucalgary.ca
    affiliation: University of Calgary
address:
  - code: University of Calgary
    address: 2500 University Drive NW, Calgary, AB
footnote:
  - code: 1
    text: "Corresponding Author"
date: "Spring 2021"
abstract: "`r paste(readLines('abstract.Rmd'), collapse = ' ')`"
keep_tex: true
keywords:
  - beetles;
  - spiders;
  - harvestmen;

@@

@@

- Precision yield data were collected directly from farmers across Alberta
  - Farmers were solicited for yield data through local agronomists, and we received data from a total of X growers across a total of X years
  - Most fields represented only a single year of data, but we did sometimes receive multiple years of data from the same field
  - File formats vary depending on the brand of combine, so we converted data to a standard format (csv) using Ag Leader SMS
  - In total, we analyzed yield data from X field-years of data, containing a total of X million data points

@@

  - To model this, we fit the following model:

  \begin{equation}
  sqrt(yield) ~ Normal (\mu, \sigma)
  \mu = Intercept + log(Polygon Size) + f(Distance from Edge_i, b=12)_i + f(Easting, Northing, b=60) + f(Harvest Sequence, b=60)
  log(\sigma) = Intercept + log(Polygon Size) + f(Distance from Edge, b=12) + f(Easting, Northing, b=60) + f(Harvest Sequence, b=60)
  \end{equation}

  - where:
  1. Polygon Size = distance traveled $\times$ width of header bar ($m^2$)
  2. Distance from Edge = distance from field edge type _i_ (m)
  3. Easting, Northing = distance from centroid of field (m)
  4. Harvest Sequence = order of harvest within field (1--N points)
  5. f(x,b) = penalized thin-plate regression spline, where x is the predictor and b is the number of basis dimensions

- In addition to modeling edge effects (our variable of interest), this model also accounts for a) differences in combine speed, b) within-field spatial variation not related to edges, and d) shifts in combine accuracy during harvest
  - Spatial or temporal variation is typically modelled using a Gaussian Process Model (e.g. Kriging) or Stochastic Partial Differential Equations (e.g. INLA), but this was computationally infeasible with 50,000 data points per field
  - Penalized splines offer a compromise, as they account for nonlinear ``wiggly" relationships in the same way as Gaussian processes but with substantially reduced computation time
  - The number of basis dimensions was checked with the _gam.check_ function from _mgcv_
  - A smoother was not used for polygon size because the relationship was easily approximated with a log-linear relationship using a single slope term
  - All models were fit in _R_ using the _mgcv_ library, and figures were created with _ggplot2_ and _ggpubr_

- Most models still had a large amount of spatio-temporal autocorrelation in residuals
  - We have 50000 data points per field, so we ignored it

- To consider results from all field-level models while also accounting for field-level uncertainty, we used an approach similar to bootstrapping of hierarchical mixed effects models
  - The approach was as follows
  1. Extract single posterior sample (_rnorm_ in R) of smoother parameters from each field-level model using coefficient estimates and standard error
  2. Use posterior sample to create new simulated smoother from each field
  3. Fit model of all simulated smoothers, and save this meta-smoother
  4. Repeat 1000 times, and calculate coverage intervals (5-95% percentiles) on saved meta-smoothers
  - This gives credible intervals for the "average" smoother while accounting for field-level variability

@@

Results here

>>>>

---
title: "Edge effects in Alberta"
author:
  - name: Samuel V. J. Robinson
    email: samuel.robinson@ucalgary.ca
    affiliation: University of Calgary
    footnote: 1
  - name: Lan H. Nguyen
    email: hoanglan.nguyen@ucalgary.ca
    affiliation: University of Calgary
  - name: Paul Galpern
    email: paul.galpern@ucalgary.ca
    affiliation: University of Calgary
address:
  - code: University of Calgary
    address: 2500 University Drive NW, Calgary, AB
footnote:
  - code: 1
    text: "Corresponding Author"
date: "Spring 2021"
abstract: "`r paste(readLines('abstract.Rmd'), collapse = ' ')`"
keep_tex: true
keywords:
  - beetles;
  - spiders;
  - harvestmen;

@@

library(tidyverse)

datSource <- read.csv('../Data/datSource.csv') #Get data source
datSourceFinal <- datSource %>% filter(use) %>% filter(grepl('Wheat|Canola',crop))
yearsPerField <- datSourceFinal %>% count(field) %>% pull(n) #How many years of data at each unique location?

@@

- Precision yield data were collected directly from farmers across Alberta
  - Farmers were solicited for yield data through local agronomists, and we received `r nrow(datSource)` field-years of data from `r length(unique(datSource$grower))` growers across a total of `r length(unique(datSource$year))` years (`r min(unique(datSource$year))`-`r max(unique(datSource$year))`)
  - We converted data to a standard csv format using Ag Leader SMS
  - `r round(100*sum(grepl('(Wheat|Canola)',datSource$crop))/nrow(datSource))`% of the crop types where either wheat (_Triticum aestivum_) or canola (_Brassica napus_), two of the most common crops in rotation in Alberta
  - The remaining crop types were poorly replicated in our sample, so we constrained our analysis to only field-years containing wheat (`r sum(datSourceFinal$crop=='Wheat')`) or canola (`r sum(datSourceFinal$crop=='Canola')`)
  - Individual fields contained between `r min(yearsPerField)` to `r max(yearsPerField)` years of data (mean: `r round(mean(yearsPerField),1)`)
  - Containing a total of `r round(sum(datSource$npoints)/1e6,1)` million data points

@@

  - To model this, we fit the following model to each field-year of data:

  \begin{equation}
  \begin{split}
  sqrt(yield) \sim & Normal (\mu, \sigma)\\
  \mu = & Intercept + log(Polygon Size) + f(Distance from Edge_i, b=12)_i + \\
   & f(Easting, Northing, b=60) + f(Sequence, b=60) \\
  log(\sigma) = & Intercept + log(Polygon Size) + f(Distance from Edge, b=12) + \\
   & f(Easting, Northing, b=60) + f(Sequence, b=60)
  \end{split}
  \end{equation}

  - where:
  1. Polygon Size = distance traveled $\times$ width of header bar ($m^2$)
  2. Distance from Edge = distance from field edge type _i_ (m)
  3. Easting, Northing = distance from centre of field (m)
  4. Sequence = order of harvest within field (1--N points)
  5. f(x,b) = penalized thin-plate regression spline, where x is the predictor and b is the number of basis dimensions

- In addition to modeling edge effects (our variable of interest), this model also accounts for a) differences in combine speed, b) within-field spatial variation not related to edges, and d) shifts in combine accuracy during harvest
  - Spatial or temporal variation is typically modelled using a Gaussian Process Model (Kriging) or approximations such as Stochastic Partial Differential Equations (e.g. INLA), but this was computationally infeasible with 50,000 data points per field
  - Penalized splines offer a compromise, as they account for nonlinear ``wiggly" relationships in the same way as Gaussian processes but with reduced dimensionality and computation time
  - The number of basis dimensions was checked with the _gam.check_ function from _mgcv_
  - The relationship between polygon size (i.e. ground speed) and yield was modeled with a log-linear relationship with a single slope term, as this closely matched smoothed versions
  - All models were fit in _R_ using the _mgcv_ library [version `r packageVersion('mgcv')`, @wood2017], and figures were created with _ggplot2_ and _ggpubr_ [versions `r packageVersion('ggplot2')`, @wickham2016; and `r packageVersion('ggpubr')`, @kassambara2020].

- To consider results from all field-level models, we fit models independent of each other, and "overall" smoothers were taken as averages of the field-level smoothers
  - However, this does not account for uncertainty in the field-level smoothers, so we used an approach similar to bootstrapping of hierarchical mixed effects models
  1. Extract single posterior sample (_rnorm_ in R) of smoother parameters from each field-level model using coefficient estimates and standard error
  2. Use posterior sample to create new simulated smoother from each field
  3. Fit new model of simulated smoothers from all fields, and save this "meta-smoother"
  4. Repeat 1000 times, and calculate coverage intervals (5-95% percentiles) on saved meta-smoothers
  - This gives coverage intervals (CIs) for the "average" smoother while accounting for field-level variability

@@

```{r , echo=FALSE, fig.cap='Fig1', out.width='100%'}  
  include_graphics('../figures/frExample.png',dpi=NA)
```

### 2021-07-10 Manuscript/manuscript.tex
Gave up on TMB
Doesn't work well enough for large numbers of samples

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage[left=3cm,right=3cm,top=3cm,bottom=3cm]{geometry}
\bibliographystyle{elsarticle-harv}
\usepackage{longtable}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Edge effects and optimal field sizes in Alberta},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

@@

  \title{Edge effects and optimal field sizes in Alberta}
    \author[University of Calgary]{Samuel V. J. Robinson\corref{1}}
   \ead{samuel.robinson@ucalgary.ca} 
    \author[University of Calgary]{Lan H. Nguyen}
   \ead{hoanglan.nguyen@ucalgary.ca} 
    \author[University of Calgary]{Paul Galpern}
   \ead{paul.galpern@ucalgary.ca} 
      \address[University of Calgary]{2500 University Drive NW, Calgary, AB}
      \cortext[1]{Corresponding Author}

@@

\begin{itemize}
\tightlist
\item
  Preserving SNL around fields is important for agriculture and conservation

  \begin{itemize}
  \tightlist
  \item
    Ecosystem services
  \item
    Act as reservoirs for beneficial insects
  \item
    Microclimate zones
  \item
    Poorly studied in North America
  \item
    Studies tend to be limited in scope (e.g.~only certain organisms/crops) and applicability
  \end{itemize}
\item
  Ecosystem services can influence both the mean and variability of yield in agroecosystems

  \begin{itemize}
  \tightlist
  \item
    Typically only averages (means) are considered, but higher stability (lower variance) in yield is also valuable
  \end{itemize}
\item
  Field size and field boundaries are directly related to the conservation of SNL in agroecosystems

  \begin{itemize}
  \tightlist
  \item
    Size examples: wheat study from the UK
  \item
    Boundary examples: flower strip studies, hedgerows
  \item
    In North America, large financial incentives to make fields large and homogeneous, especially with large harvesting and planting equipment
  \item
    Crop edge effects cause low yields at the margins of fields because of late emergence, poor microclimate, and competition with weeds
  \item
    Ecosystem services should decay with distance from edge, so crops at the centre of a large field will not benefit from ecosystem services
  \item
    Therefore, there should be a ``goldilocks" field size, where negative edge effects are canceled out by ecosystem services
  \end{itemize}
\item
  Precision yield data holds enormous promise for agronomy

  \begin{itemize}
  \tightlist
  \item
    Limited because of:
  \item
    Lack of standardized formats
  \item
    Yearly calibration data
  \item
    Clear statistical protocols
  \end{itemize}
\end{itemize}

@@

  \begin{itemize}
  \tightlist
  \item
    Farmers were solicited for yield data through local agronomists, and we received data from a total of X growers across a total of X years
  \item
    Most fields represented only a single year of data, but we did sometimes receive multiple years of data from the same field
  \item
    File formats vary depending on the brand of combine, so we converted data to a standard format (csv) using Ag Leader SMS
  \item
    In total, we analyzed yield data from X field-years of data, containing a total of X million data points
  \end{itemize}
\item
  Yield data was collected in discrete rectangles of the same length as the data interval (distance = speed \(\times\) interval, typically 1 second) and the same width as the combine header (5-7 m)

  \begin{itemize}
  \tightlist
  \item
    We extracted the size of each polygon (m\(^2\)), dry yield (tonnes), and the spatial location, and the sequence of collection (1 - end of harvest)
  \item
    Because of the large number of yield rectangles per field (30-800 thousand), we used the centroid of each polygon as its location
  \item
    Seeding and application rates were constant with each field, so we did not consider inputs in our analysis
  \end{itemize}
\item
  Field boundaries were automatically digitized, then manually checked using satellite imagery and classified land cover data

  \begin{itemize}
  \tightlist
  \item
    Crop boundaries are flexible, and often change from year to year depending on planting and emergence conditions (e.g.~flooding during some years)
  \item
    Additionally, seminatural features often change from year to year

    \begin{itemize}
    \tightlist
    \item
      Ephemeral wetlands are flooded during some years, but consist mainly of grasses during dry years
    \item
      Grass boundaries can change if fields are used for as haying or pasture during crop rotation
    \end{itemize}
  \item
    This makes accurate and consistent classification of field boundaries very difficult
  \item
    Because of this, we used the following general categories for field boundaries:

    \begin{enumerate}
    \def\labelenumi{\arabic{enumi}.}
    \tightlist
    \item
      Standard: thin (\(<10\) m wide) grassy field boundary, often grassy road right-of-ways
    \item
      Wetland: permanent wetland, whose borders are largely unchanged from year-to-year
    \item
      Forest/shrub: permanent windbreaks or remnant forests
    \item
      Grass: larger grassy area (pasture or permanent seminatural grassland)
    \item
      Other crop: different crop with little or no visible boundary between planted areas
    \item
      Bare: unplanted (fallow), flooded area, or unpaved roads
    \end{enumerate}
  \end{itemize}
\end{itemize}

@@

\begin{itemize}
\tightlist
\item
  At each field, we fit an additive model of the effect of boundary distance on crop yield while accounting for within-field spatial variation and temporal variation in the combine yield monitor

  \begin{itemize}
  \tightlist
  \item
    Crop yield varies within a field due to soil conditions, moisture, seeding rates, herbicide application, and previous agricultural practices such as strip farming
  \item
    While sensor calibration can reduce combine-level bias (such as a combine recording consistently higher/lower yields), this does not address sensor drift over time that occurs within fields
  \item
    The yield monitors may record lower yields as sensors accumulate debris during harvest (pers. comm. Trent Clark), leading to changes in accuracy and bias over time
  \item
    Additionally, ground speed is known to be extremely important to yield monitor accuracy (Arslan \& Colvin 2002)
  \item
    To address this, we fit the following model:
  \end{itemize}

  \begin{equation}
  ln(yield) 
  \end{equation}
\end{itemize}

@@

>>>>

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage[left=3cm,right=3cm,top=3cm,bottom=3cm]{geometry}
\bibliographystyle{elsarticle-harv}
\usepackage{longtable}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Edge effects in Alberta},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

@@

  \title{Edge effects in Alberta}
    \author[University of Calgary]{Samuel V. J. Robinson\corref{1}}
   \ead{samuel.robinson@ucalgary.ca} 
    \author[University of Calgary]{Lan H. Nguyen}
   \ead{hoanglan.nguyen@ucalgary.ca} 
    \author[University of Calgary]{Paul Galpern}
   \ead{paul.galpern@ucalgary.ca} 
      \address[University of Calgary]{2500 University Drive NW, Calgary, AB}
      \cortext[1]{Corresponding Author}

@@

\begin{itemize}
\tightlist
\item
  Intensive agricultural production has increased over the last 100 years, and agricultural land now makes up over a third of ice-free land on Earth

  \begin{itemize}
  \tightlist
  \item
    This has allows increases in human population and increased (global) stability in production
  \item
    However, this is not without cost, as higher-diversity non-crops are converted to lower-diversity crops, resulting in loss of habitat and overall biodiversity of non-target organisms
  \item
    Maintaining both biodiversity and production in agroecosystems represents a seldom-considered goal of conservationists and agronomists, and hold the potential for win-win scenarios
  \item
    Key to this is the preservation of semi-natural land (SNL), which represents the interface between crops and non-crops within agroecosystems
  \end{itemize}
\item
  SNL in and around crops is important for both agricultural production and conservation

  \begin{itemize}
  \tightlist
  \item
    They are habitat for mobile organisms, and can therefore act as sources of ecosystem services such as pollination or pest control
  \item
    They also can create microclimate effects that reduce extreme temperature, trap moisture, and reduce wind speed
  \item
    Unfortunately, most of the research is concentrated in Europe, and they tend to be less-studied on other continents
  \item
    In particular, North American agroecosystems have larger fields, and different varieties and agronomic practices, all of which could negate effects of SNL
  \end{itemize}
\end{itemize}

\begin{itemize}
\tightlist
\item
  SNL may affect yields at intermediate distances, depending on the spatial scale at which ecosystem services operate

  \begin{itemize}
  \tightlist
  \item
    Edge effects cause low yields at the edge of crops because of sparse or late seedling emergence, poor microclimate, and competition with weeds
  \item
    At the same time, the centre of large fields will not receive ecosystem services if they decay with distance from edges
  \item
    For example, pollination services from central place foragers that nest in SNL but forage in crops drops rapidly with distance
  \item
    Therefore, yield may be maximized at intermediate distances, where the ecosystem services cancel out negative edge effects
  \item
    This suggests a ``goldilocks" field size, where negative edge effects are canceled out by ecosystem services
  \end{itemize}
\end{itemize}

\begin{itemize}
\tightlist
\item
  Studies of SNL effects on crop yield also suffer from limited scope (e.g.~few crop types) and small sample sizes, limiting inference and reducing generality (but see wheat study)

  \begin{itemize}
  \tightlist
  \item
    For this reason, large-scale precision yield data holds enormous promise for agronomy, as it allows
  \item
    However, its use is limited for several reasons: 1) Lack of standardized formats between equipment types, 2) Sensor calibration required for field-level accuracy, and 3) unfamiliarity with spatial statistics
  \item
    Ecosystem services can influence both the mean and variability of yield in agroecosystems
  \item
    Typically only averages (means) are considered, but higher stability (lower variance) in yield can also be valuable
  \item
    Size examples: wheat study from the UK
  \item
    There are few studies of yield variability (only those with large datasets), but precision yield data opens up the possibility of modeling within-field variability, as well as average yield
  \end{itemize}
\item
  In this paper we ask:

  \begin{enumerate}
  \def\labelenumi{\arabic{enumi}.}
  \tightlist
  \item
    How does crop yield change with distance from the edge of field?
  \item
    Does this depend on type of field edge?
  \item
    Is there an intermediate distance where yield is maximized or variance is minimized?
  \end{enumerate}
\end{itemize}

@@

  \begin{itemize}
  \tightlist
  \item
    Farmers were solicited for yield data through local agronomists, and we received 298 field-years of data from 5 growers across a total of 7 years (2014-2020)
  \item
    We converted data to a standard csv format using Ag Leader SMS
  \item
    72\% of the crop types where either wheat (\emph{Triticum aestivum}) or canola (\emph{Brassica napus}), two of the most common crops in rotation in Alberta
  \item
    The remaining crop types were poorly replicated in our sample, so we constrained our analysis to only field-years containing wheat (94) or canola (119)
  \item
    Individual fields contained between 1 to 5 years of data (mean: 2.7)
  \item
    Containing a total of 18.4 million data points
  \end{itemize}
\item
  Yield data is collected in rectangles of the same length as the data interval (distance = combine ground speed \(\times\) interval, typically 1 second) and the same width as the combine header (5-7 m)

  \begin{itemize}
  \tightlist
  \item
    We extracted the size of each polygon (m\(^2\)), dry yield (tonnes), and the spatial location, and the sequence of collection (1 - end of harvest)
  \item
    Because of the large number of yield rectangles per field (30-800 thousand), we used the centroid of each polygon as its location, treating areal data as point data
  \item
    Seeding and application rates were constant across fields, so we did not consider inputs in our analysis
  \item
    We used dry yield (tonnes of seed/hectare after accounting for crop moisture) as our measure of crop yield
  \item
    Due to the large number of data at each field, we sub-sampled to 50,000 data points per field to reduce computation time
  \end{itemize}
\item
  Field boundaries were automatically digitized using buffers from the yield data locations, then manually checked using satellite imagery from Google Earth and classified land cover data from AAFC

  \begin{itemize}
  \tightlist
  \item
    Crop boundaries are flexible, and often change yearly depending on planting and emergence conditions (e.g.~flooding during some years)
  \item
    Additionally, seminatural features often change yearly

    \begin{itemize}
    \tightlist
    \item
      Ephemeral wetlands are flooded during some years, but consist mainly of grasses during dry years
    \item
      Grass boundaries can change if fields are used for as haying or pasture during crop rotation
    \end{itemize}
  \item
    This makes accurate and consistent classification of field boundaries difficult
  \item
    We defined the following general categories for field boundaries:

    \begin{enumerate}
    \def\labelenumi{\arabic{enumi}.}
    \tightlist
    \item
      Standard: grassy field edge, staging yard, or road right-of-way (grassy strip typically 5-10m wide)
    \item
      Wetland: permanent wetland; borders are largely unchanged from year-to-year
    \item
      Shelterbelt: permanent windbreaks, shelterbelts, remnant forests, or shrublands
    \item
      Other crop: annual crop or pasture with little or no visible boundary between planted areas
    \item
      Bare: unplanted, fallow, flooded area, temporary wetland (only present for a single season), staging yard, oil and gas equipment, or road without a planted boundary
    \item
      Grassland: permanent seminatural grassland or pasture (not in rotation)
    \end{enumerate}
  \end{itemize}
\end{itemize}

@@

\begin{itemize}
\item
  At each field, we fit an additive model of the effect of boundary distance on crop yield while accounting for within-field spatial variation and temporal variation in the combine yield monitor

  \begin{itemize}
  \item
    Crop yield can vary within a field due to soil conditions, moisture, seeding rates, herbicide application, and previous agricultural practices
  \item
    Ground speed is extremely important to yield monitor accuracy (Arslan \& Colvin 2002), with low ground speed registering higher yields
  \item
    While sensor calibration can reduce combine-level bias (such as a combine recording consistently higher/lower yields across fields), this does not address sensor drift that occurs over time within fields
  \item
    This may be caused by sensors accumulating debris during harvest (pers. comm. Trent Clark), leading to changes in accuracy and bias over time
  \item
    To model this, we fit the following model to each field-year of data:
  \end{itemize}

  \begin{equation}
  \begin{split}
  sqrt(yield) \sim & Normal (\mu, \sigma)\\
  \mu = & Intercept + log(Polygon Size) + f(Distance from Edge_i, b=12)_i + \\
   & f(Easting, Northing, b=60) + f(Sequence, b=60) \\
  log(\sigma) = & Intercept + log(Polygon Size) + f(Distance from Edge, b=12) + \\
   & f(Easting, Northing, b=60) + f(Sequence, b=60)
  \end{split}
  \end{equation}

  \begin{itemize}
  \tightlist
  \item
    where:
  \end{itemize}

  \begin{enumerate}
  \def\labelenumi{\arabic{enumi}.}
  \tightlist
  \item
    Polygon Size = distance traveled \(\times\) width of header bar (\(m^2\))
  \item
    Distance from Edge = distance from field edge type \emph{i} (m)
  \item
    Easting, Northing = distance from centre of field (m)
  \item
    Sequence = order of harvest within field (1--N points)
  \item
    f(x,b) = penalized thin-plate regression spline, where x is the predictor and b is the number of basis dimensions
  \end{enumerate}
\item
  In addition to modeling edge effects (our variable of interest), this model also accounts for a) differences in combine speed, b) within-field spatial variation not related to edges, and d) shifts in combine accuracy during harvest

  \begin{itemize}
  \tightlist
  \item
    Spatial or temporal variation is typically modelled using a Gaussian Process Model (Kriging) or approximations such as Stochastic Partial Differential Equations (e.g.~INLA), but this was computationally infeasible with 50,000 data points per field
  \item
    Penalized splines offer a compromise, as they account for nonlinear ``wiggly" relationships in the same way as Gaussian processes but with substantially reduced computation time
  \item
    The number of basis dimensions was checked with the \emph{gam.check} function from \emph{mgcv}
  \item
    The relationship between polygon size (i.e.~ground speed) and yield was modeled with a log-linear relationship with a single slope term, as this closely matched smoothed versions
  \item
    All models were fit in \emph{R} using the \emph{mgcv} library (version 1.8.36, Wood 2017), and figures were created with \emph{ggplot2} and \emph{ggpubr} (versions 3.3.3, Wickham 2016; and 0.4.0, Kassambara 2020).
  \end{itemize}
\item
  To consider results from all field-level models, we fit models independent of each other, and ``overall'' smoothers were taken as averages of the field-level smoothers

  \begin{itemize}
  \tightlist
  \item
    However, this does not account for uncertainty in the field-level smoothers, so we used an approach similar to bootstrapping of hierarchical mixed effects models
  \end{itemize}

  \begin{enumerate}
  \def\labelenumi{\arabic{enumi}.}
  \tightlist
  \item
    Extract single posterior sample (\emph{rnorm} in R) of smoother parameters from each field-level model using coefficient estimates and standard error
  \item
    Use posterior sample to create new simulated smoother from each field
  \item
    Fit new model of simulated smoothers from all fields, and save this ``meta-smoother''
  \item
    Repeat 1000 times, and calculate coverage intervals (5-95\% percentiles) on saved meta-smoothers
  \end{enumerate}

  \begin{itemize}
  \tightlist
  \item
    This gives coverage intervals (CIs) for the ``average'' smoother while accounting for field-level variability
  \end{itemize}
\end{itemize}

@@

\leavevmode\hypertarget{ref-kassambara2020}{}%
Kassambara, A. (2020). \emph{{ggpubr}: 'ggplot2' based publication ready plots}.

\leavevmode\hypertarget{ref-wickham2016}{}%
Wickham, H. (2016). \emph{{ggplot2}: Elegant graphics for data analysis}. Springer-Verlag New York.

\leavevmode\hypertarget{ref-wood2017}{}%
Wood, S.N. (2017). \emph{Generalized additive models: An introduction with {R}}. CRC press.
